---
title: "Multiple imputation"
author: "Rianne"
date: "2026-03-18"
output: html_document
---
#Survival packages 
```{r}
library(ggsurvfit)
library(survival)
library(survminer)
library(ggplot2)
```

#Loading in the data
```{r}
library(haven)
data <- read_sav("t821_martanew.sav")
subdata <- data[, c('gender','age','wbccat','anthra_dose','relapse','secmalign',
                    'death','time_surv',
                    'event','time_event',
                    'event_CR','time_CR')]
set.seed(42)
```



#Multiple Imputation: Filling in mean, max and min
Of course we need the filled in value for complete remission to be lower than the survival time
```{r}
mean_tCR <- mean(subdata$time_CR[which(!is.na(subdata$time_CR))])
print("For which missing data will the overall mean not work:")
which(subdata$time_surv[which(is.na(subdata$time_CR) & subdata$event_CR == 0)] <= mean_tCR )
min_tCR <- min(subdata$time_CR[which(!is.na(subdata$time_CR))])
print("For which missing data will the overall min not work:")
which(subdata$time_surv[which(is.na(subdata$time_CR) & subdata$event_CR == 0)] <= min_tCR )
max_tCR <- max(subdata$time_CR[which(!is.na(subdata$time_CR))])
print("For which missing data will the overall max not work:")
which(subdata$time_surv[which(is.na(subdata$time_CR) & subdata$event_CR == 0)] <= max_tCR )
#rm(list = c("mean_tCR", "min_tCR", "max_tCR"))
```
Notice that this does not work for the maximum so we need a better way to fill it in than the overall mean, minimum and maximum.


```{r}
#The filled in mean vector
time_CR_meanfill <- subdata$time_CR #Vector to be filled
missing <- which(is.na(subdata$time_CR) & data$event_CR == 0)#Which data is missing
for (value in missing) {
  time_CR_meanfill[value] = 
    mean(subdata$time_CR[which(!is.na(subdata$time_CR))])
}

#The filled in min vector
time_CR_minfill <- subdata$time_CR #Vector to be filled
missing <- which(is.na(subdata$time_CR) & data$event_CR == 0)#Which data is missing
for (value in missing) {
  time_CR_minfill[value] = 
    min(subdata$time_CR[which(!is.na(subdata$time_CR))])
}

#The filled in max vector
time_CR_event_maxfill <- subdata$time_CR #Vector to be filled for event
time_CR_death_maxfill <- subdata$time_CR #Vector to be filled for death
missing <- which(is.na(subdata$time_CR) & data$event_CR == 0)#Which data is missing
for (value in missing) {
  time_CR_event_maxfill[value] = 
    max(subdata$time_CR[which(subdata$time_CR <= subdata$time_event[value] 
                        & !is.na(subdata$time_CR))])
  time_CR_death_maxfill[value] = 
    max(subdata$time_CR[which(subdata$time_CR <= subdata$time_surv[value] 
                        & !is.na(subdata$time_CR))])
}
```


##Plotting estimators time since complete remission
```{r}
#Making a subset and 8 new columns we need
subdata_CR_fill <- subdata#New dataset 

subdata_CR_fill$time_event_sinceCR <- #event 
  subdata_CR_fill$time_event - subdata_CR_fill$time_CR
subdata_CR_fill$time_surv_sinceCR <- #death
  subdata_CR_fill$time_surv - subdata_CR_fill$time_CR

subdata_CR_fill$time_event_sinceCR_meanfill <- #event and mean
  subdata_CR_fill$time_event - time_CR_meanfill
subdata_CR_fill$time_surv_sinceCR_meanfill <- #death and mean
  subdata_CR_fill$time_surv - time_CR_meanfill

subdata_CR_fill$time_event_sinceCR_minfill <- #event and min
  subdata_CR_fill$time_event - time_CR_minfill
subdata_CR_fill$time_surv_sinceCR_minfill <- #death and min
  subdata_CR_fill$time_surv - time_CR_minfill

subdata_CR_fill$time_event_sinceCR_maxfill <- #event and max
  subdata_CR_fill$time_event - time_CR_event_maxfill
subdata_CR_fill$time_surv_sinceCR_maxfill <- #death and max
  subdata_CR_fill$time_surv - time_CR_death_maxfill
```



```{r}
#Plotting EFS
##original
fit_event_km_CR <- survfit(Surv(time_event_sinceCR, event) ~ 1, 
                              data = subdata_CR_fill)#Kaplan-Meier
df_event_km_CR <- data.frame(time = fit_event_km_CR$time,
                          surv = fit_event_km_CR$surv,
                          Estimator = "Known values")
##mean
fit_event_km_CR_meanfill <- survfit(Surv(time_event_sinceCR_meanfill, event) ~ 1, 
                              data = subdata_CR_fill)#Kaplan-Meier
df_event_km_CR_meanfill <- data.frame(time = fit_event_km_CR_meanfill$time,
                          surv = fit_event_km_CR_meanfill$surv,
                          Estimator = "Known values and missing values are filled with the mean")
##min
fit_event_km_CR_minfill <- survfit(Surv(time_event_sinceCR_minfill, event) ~ 1, 
                              data = subdata_CR_fill)#Kaplan-Meier
df_event_km_CR_minfill <- data.frame(time = fit_event_km_CR_minfill$time,
                          surv = fit_event_km_CR_minfill$surv,
                          Estimator = "Known values and missing values are filled with the minimum")
##max
fit_event_km_CR_maxfill <- survfit(Surv(time_event_sinceCR_maxfill, event) ~ 1, 
                              data = subdata_CR_fill)#Kaplan-Meier
df_event_km_CR_maxfill <- data.frame(time = fit_event_km_CR_maxfill$time,
                          surv = fit_event_km_CR_maxfill$surv,
                          Estimator = "Known values and missing values are filled with the maximum")

df_event_CR_fill <- rbind(df_event_km_CR, df_event_km_CR_meanfill,
                          df_event_km_CR_minfill, df_event_km_CR_maxfill )#combining the dataframes

plot_event_CR_fill <- ggplot(df_event_CR_fill, aes(time, surv, color = Estimator)) +
  geom_step()+
  labs(
    title = "Event Free Survival using the Kaplan-Meier estimator",
    x = "Time since complete remission (in years)",
    y = "Survival function"
    )+
  xlim(0,10)+
  ylim(0,1)+
  theme(legend.position = c(0.3, 0.25),
        legend.background = element_rect(fill = "white"))
print(plot_event_CR_fill)
ggsave("Event_free_survival_filled.png", plot = plot_event_CR_fill, 
       width = 6.5, height = 4, dpi = 300)

#Plotting OS
##original
fit_death_km_CR <- survfit(Surv(time_surv_sinceCR, death) ~ 1, 
                              data = subdata_CR_fill)#Kaplan-Meier
df_death_km_CR <- data.frame(time = fit_death_km_CR$time,
                          surv = fit_death_km_CR$surv,
                          Estimator = "Known values")
##mean
fit_death_km_CR_meanfill <- survfit(Surv(time_surv_sinceCR_meanfill, death) ~ 1, 
                              data = subdata_CR_fill)#Kaplan-Meier
df_death_km_CR_meanfill <- data.frame(time = fit_death_km_CR_meanfill$time,
                          surv = fit_death_km_CR_meanfill$surv,
                          Estimator = "Known values and missing values are filled with the mean")
##min
fit_death_km_CR_minfill <- survfit(Surv(time_surv_sinceCR_minfill, death) ~ 1, 
                              data = subdata_CR_fill)#Kaplan-Meier
df_death_km_CR_minfill <- data.frame(time = fit_death_km_CR_minfill$time,
                          surv = fit_death_km_CR_minfill$surv,
                          Estimator = "Known values and missing values are filled with the minimum")
##max
fit_death_km_CR_maxfill <- survfit(Surv(time_surv_sinceCR_maxfill, death) ~ 1, 
                              data = subdata_CR_fill)#Kaplan-Meier
df_death_km_CR_maxfill <- data.frame(time = fit_death_km_CR_maxfill$time,
                          surv = fit_death_km_CR_maxfill$surv,
                          Estimator = "Known values and missing values are filled with the maximum")

df_death_CR_fill <- rbind(df_death_km_CR, df_death_km_CR_meanfill,
                          df_death_km_CR_minfill, df_death_km_CR_maxfill )#combining the dataframes

plot_death_CR_fill <- ggplot(df_death_CR_fill, aes(time, surv, color = Estimator)) +
  geom_step()+
  labs(
    title = "Overall Survival using the Kaplan-Meier estimator",
    x = "Time since complete remission (in years)",
    y = "Survival function"
    )+
  xlim(0,10)+
  ylim(0,1)+
  theme(legend.position = c(0.3, 0.25),
        legend.background = element_rect(fill = "white"))
print(plot_death_CR_fill)
ggsave("Overall_survival_filled.png", plot = plot_death_CR_fill, 
       width = 6.5, height = 4, dpi = 300)
```

#To show that the event and survival times of the missing data are just lower
```{r}
mean(subdata$time_event[which(!is.na(subdata$time_CR)&!is.na(subdata$time_event))])#Observed
mean(subdata$time_event[which(is.na(subdata$time_CR)&subdata$event_CR == 0)])#Missing

mean(subdata$time_surv[which(!is.na(subdata$time_CR)&!is.na(subdata$time_surv))])#Observed
mean(subdata$time_surv[which(is.na(subdata$time_CR)&subdata$event_CR == 0)])#Missing
```















