---
title: "Tables"
author: "Rianne"
date: "2026-02-24"
output: github_document
---

#Loading in the data
```{r}
library(haven)
data <- read_sav("t821_martanew.sav")
subdata <- data[, c('gender','age','wbccat','anthra_dose','relapse','secmalign',
                    'death','time_surv',
                    'event','time_event',
                    'event_CR','time_CR')]
```

#Summary of the interesting data
```{r}
summary(subdata)
str(subdata)
```

##relapse, death, relapse and death, event, event and death
```{r}
print("death:")
table(subdata$death)
cat("\n")
print("relapse:")
table(subdata$relapse)
cat("\n")
print("relapse vertical against horizontal death:")
table(subdata$relapse,subdata$death)
cat("\n")
print("event:")
table(subdata$event)
cat("\n")
print("missing data:")
table(subdata$event_CR[is.na(subdata$time_CR)])
```

#Reason why imputation function are steeper and lower

```{r}
table(data$event[which(!is.na(data$time_CR) & data$event_CR == 0)])
172/488
table(data$event[which(is.na(data$time_CR) & data$event_CR == 0)])
117/284
table(data$death[which(!is.na(data$time_CR) & data$event_CR == 0)])
111/488
table(data$death[which(is.na(data$time_CR) & data$event_CR == 0)])
81/284
```


#Baseline characteristics of the patients

##Baseline gender of the patients
```{r}
table(subdata$gender)#1 is male
```
##Baseline age of the patients
```{r}
cat("Age less than 6 years:", length(which(subdata$age < 6)),"\n")
cat("Age 6 to less than 9 years:", length(which(subdata$age >= 6 & subdata$age < 9)),"\n")
cat("Age 9 to less than 13 years:", length(which(subdata$age >= 9 & subdata$age < 13)),"\n")
cat("Age 13 years or more:", length(which(subdata$age >= 13)),"\n")
```
##Baseline white blood cells of the patients
```{r}
table(subdata$wbccat) #"<10", "10-<20", "20-<50", "50-<100", ">100", "missing"
```
## Baseline anthracyclines dose in induction of the patients
```{r}
cat("Anthra_dose of less than 300:", length(which(subdata$anthra_dose < 300)),"\n")
cat("Anthra_dose of 300 to less than 360:", length(which(subdata$anthra_dose >= 300 & 
                                                          subdata$anthra_dose < 360)),"\n")
cat("Anthra_dose of 360 to less than 420:", length(which(subdata$anthra_dose >= 360 & 
                                                           subdata$anthra_dose < 420)),"\n")
cat("Anthra_dose of 420 or more:", length(which(subdata$anthra_dose >= 420)),"\n")
```

